---
title: "Range tests for equality and equivalence across specifications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Range tests for equality and equivalence across specifications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(robustness)
```

## The method

When an empirical finding is reported across many specifications, readers want
to know whether the estimates agree. Two questions are distinct. *Equality*
asks whether the specifications could share a common value. *Equivalence* asks
whether they are close enough that any differences are immaterial. Jaeger
(2026) builds both tests on the range of the estimates,
$R(\hat{\boldsymbol\theta}) = \max_k \hat\theta_k - \min_k \hat\theta_k$, and on
a companion Wald statistic, using the bootstrap to obtain their distributions.

This package computes those statistics. It takes the full-sample estimates and
a matrix of bootstrap draws and returns, for each comparison, the range and
Wald statistics, equivalence bounds $\delta^*(\alpha)$, and equality-test
p-values. It is the R counterpart of the Stata `robustness` command; given the
same draws, the two produce the same numbers.

The package computes statistics from draws. It does not generate the draws. The
section below shows how.

## A first example

The bundled `lee_example` is a small illustration shaped like the Lee (2008)
incumbency application: seven specifications of an effect near 0.077.

```{r}
str(lee_example, max.level = 1)

robustness(lee_example$theta, lee_example$draws,
           comparisons = lee_example$comparisons)
```

Comparisons are given as a named list of column indices, defined directly in R.
This keeps them transparent and self-documenting: each comparison is the set of
specifications it pools.

The result has a tidy form for assembling tables:

```{r}
res <- robustness(lee_example$theta, lee_example$draws,
                  comparisons = lee_example$comparisons)
as.data.frame(res)
```

## Generating the draws

The statistics need uncentred bootstrap draws of the coefficient, one column
per specification. The essential discipline is the **shared resample**: on each
bootstrap replication, draw the resample *once* and estimate every
specification on that same resample. Resampling separately for each
specification destroys the joint distribution and silently gives wrong
p-values and bounds.

Here is a minimal generation example in R, in the spirit of the Lee
application: a clustered bootstrap over a single resample per replication.

```{r, eval = FALSE}
# `dat` has outcome y, running variable x, cluster id, and a treatment d.
# Seven specifications differ in controls; all share the resample.

specs <- list(
  no_controls  = y ~ d,
  prev_vote    = y ~ d + lag_vote,
  office_exp   = y ~ d + office,
  elect_exp    = y ~ d + elect,
  all_controls = y ~ d + lag_vote + office + elect,
  resid_out    = yresid ~ d,
  first_diff   = dy ~ d
)
theta <- sapply(specs, function(f) coef(lm(f, data = dat))[["d"]])

clusters <- unique(dat$cluster)
B <- 9999
draws <- matrix(NA_real_, B, length(specs))
set.seed(42)
for (b in seq_len(B)) {
  # ONE resample of clusters, reused for every specification
  pick <- sample(clusters, length(clusters), replace = TRUE)
  bdat <- do.call(rbind, lapply(pick, function(cl) dat[dat$cluster == cl, ]))
  draws[b, ] <- sapply(specs, function(f) coef(lm(f, data = bdat))[["d"]])
}

robustness(theta, draws, comparisons = list(all = 1:7, main = 1:5))
```

The single `sample()` call per replication, reused across all specifications,
is the shared resample. That one line is what makes the joint distribution
correct.

## Using draws generated in Stata

The Stata `robustness` package saves draws as `.dta` files. To analyse them in
R, read them with `haven` and pass the pieces to `robustness()`. The conversion
is a few lines and lives in your own script, not in this package.

```{r, eval = FALSE}
library(haven)

meta  <- read_dta("output/draws/lee_draws_meta.dta")
draws <- read_dta("output/draws/lee_draws.dta")

theta <- meta$theta
D     <- as.matrix(draws[, grep("^coef[0-9]+$", names(draws))])

robustness(theta, D,
           comparisons = list(all = 1:7, main = 1:5,
                              main_resid = 1:6,
                              main_firstdiff = c(1, 2, 3, 4, 5, 7)))
```

Because R and Stata use different random number generators, draws generated
separately in the two languages are not identical. At a realistic number of
replications the resulting statistics agree up to Monte Carlo error: the test
is designed not to depend on the particular draws. Reading the *same* saved
Stata draws into R, as above, reproduces the Stata statistics exactly.

## Citing

To cite the method, cite the paper:

```{r, eval = FALSE}
citation("robustness")
```

## Reference

Jaeger, David A. (2026). Robustness? Range Tests for Equality and Equivalence
Across Specifications.
